Really Hard Instances for the Subgraph Isomorphism Problem

### 2015-09-16 paper/paper.tex
Bit more

\begin{abstract}
    Really hard SIP instances.
\end{abstract}

@@

Here we present and evaluate a new method for creating random pattern/target pairs which can
generate both satisfiable and unsatisfiable instances. Our work builds upon the widely studied phase
transition phenomena observed in satisfiability and graph colouring problems. However, we have three
control parameters rather than one: we can independently alter the density of the pattern graph, the
density of the target graph, and the relative orders (number of vertices) of the pattern and target
graphs.  For non-induced isomorphisms, with the correct choice of parameters there is a phase
transition from satisfiable to unsatisfiable, and as expected we see a complexity peak occur (with
three different solvers) near this phase transition.

For certain choices of parameter for induced isomorphisms, there are two phase transitions, going
from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
really hard to solve. However, each of the three solvers we tested also finds the central
unsatisfiable region to be really hard---this is surprising, since this region is apparently
over-constrained. To show that this is not a simple weakness of subgraph isomorphism algorithms, we
show that the same behaviour occurs when using use a SAT solver, a pseudo-boolean solver, and a
clique encoding.

@@

?? experimental setup.

@@

\begin{figure*}
    \input{gen-graph-induced.tex}
    \setlength{\abovecaptionskip}{-2em}
    \caption{Behaviour of algorithms on the induced variant. Each point is the average of ten
        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
        satisfiable; the white bands shows the phase transitions. On the second row, we show the
        number of search nodes used by the Glasgow algorithm, and on the third row, the number of
        search nodes used by the VF2 algorithm: the dark regions indicate ``really hard''
        instances.}\label{figure:induced}
\end{figure*}

\section{Other Encodings}

@@

>>>>

\begin{abstract}
    Really hard SIP instances. Phase transitions. Alternative models and reduction.
\end{abstract}

@@

Here we present and evaluate a new method for creating random pattern/target pairs which can
generate both satisfiable and unsatisfiable instances. Our work builds upon the widely studied phase
transition phenomena observed in satisfiability and graph colouring problems ?? cite stuff. However,
we have three control parameters rather than one: we can independently alter the density of the
pattern graph, the density of the target graph, and the relative orders (number of vertices) of the
pattern and target graphs.  For non-induced isomorphisms, with the correct choice of parameters
there is a phase transition from satisfiable to unsatisfiable, and as expected we see a complexity
peak occur (with three different solvers) near this phase transition.

For certain choices of parameter for induced isomorphisms, there are two phase transitions, going
from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
really hard to solve. However, each of the three solvers we tested also finds the central
unsatisfiable region to be really hard---this is surprising, since this region is apparently
over-constrained. To show that this is not a simple weakness of subgraph isomorphism algorithms, we
verify that the same behaviour occurs when using use a SAT solver, a pseudo-boolean solver, and a
clique encoding. ?? Except maybe it doesn't...

@@

?? Software versions.

Our experiments were performed on systems with Intel Xeon E5-4650 v2 (Q1'14) CPUs and 768GBytes RAM
(although this much RAM was not needed), running Scientific Linux release 6.7. Software was compiled
using GCC 4.9.

@@

\begin{figure*}
    \hskip-1.2cm\input{gen-graph-induced.tex}
    \setlength{\abovecaptionskip}{-2em}
    \caption{Behaviour of algorithms on the induced variant. Each point is the average of ten
        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
        satisfiable; the white bands shows the phase transitions. On the second row, we show the
        number of search nodes used by the Glasgow algorithm, and on the third row, the number of
        search nodes used by the VF2 algorithm: the dark regions indicate ``really hard''
        instances.}\label{figure:induced}
\end{figure*}

\subsection{Predicting Induced Behaviour}

?? Graph showing the product of satisfiable with vertical flip, min (unsat) and max (sat) of search
nodes with vertical flip.

\section{Other Encodings and Solvers}

@@

?? \cite{Anton:2009}

?? \cite{Lipets:2009}
